# Section 3
### Data science question 1: 
What is the distribution of country counts across the range of income classifications? 

This question is examining the distribution of countries across the income classifications, which range from 0-3. This question seeks to determine which income classification has the highest number of countries, and which income classification has the lowest. 

My data analysis method starts with wrangling the data to only include each country as one observation (in the original dataframe, most countries are listed multiple times due to several different observation years). Because income classification remains the same across all these years for each country, I was able to simply remove duplicates of country names so that each country was only included as one observation. Next, I made the bar plot of counts using ggplot by including the built-in “stat” argument to have the height of the bar represent the count of cases within each income classification. 

```{r part_2, echo=FALSE}
plot(income_classification_graph)
```

Evaluation: 
The bar graph reveals that income classification “2” has the highest country count, around approximately 55, and income classification “3” has the lowest country count, at just above 20. It appears that the majority of countries lie in the middle of the distribution of income classifications, across classification “1” and “2”. 

### Data science question 2: 
What is the distribution of countries across the range of health expenditures? Does this distribution map onto the distribution of income classifications? 

This question is examining the distribution of countries across the range of health expenditures, which range from 102 to 994. This question further seeks to examine whether this distribution correlates to the distribution of income classifications. 

My data analysis method starts with wrangling the data to join our two datasets together in order to evaluate both the income classification and the health expenditures in the same dataset. I used the health expenditure data from 2018 because that was the most recent year with data. I arranged the dataset by income classification so that it follows the same order as the income classification graph, and made the bar graph using the stat function again, very similarly to the income classification graph. 

```{r part_2, echo=FALSE}
plot(health_expenditures_graph)
```

Evaluation: 
The bar graph reveals that the distribution of health expenditures varies a good amount. It looks like the health expenditure amount with the highest frequency of countries is 994, which also happens to be the highest health expenditure amount. This is fairly surprising when compared to the income classification distribution, since income classification 3 has the lowest frequency of countries. You would expect that primarily the countries in income classification 3 would have higher health expenditures, so the distributions would look similar. However, this distribution of health expenditures does not seem to have a distinct pattern that correlates with the income classification distribution, and looks pretty evenly dispersed. This indicates that there is not a correlation between income classification and health expenditures. 
