# Section 2
The first dataset we will be using is a dataset of all the health expenditures per capita in each country. The second dataset we will be using includes the statistics on malnutrition amongst children aged 0-59 months, broken down by the four types of malnutrition: wasting, stunting, underweight, and overweight.

The first dataset comes from data prepared by the World Health Organization. “The health expenditure estimates have been prepared by the World Health Organization under the framework of the System of Health Accounts 2011 (SHA 2011). The Health SHA 2011 tracks all health spending in a given country over a defined period of time regardless of the entity or institution that financed and managed that spending. It generates consistent and comprehensive data on health spending in a country, which in turn can contribute to evidence-based policy-making.” (World Development Indicators)
Link: https://databank.worldbank.org/reports.aspx?source=2&series=SH.XPD.CHEX.PC.CD

The second dataset comes from data compiled by Ruchi Bhatia, originally sourced from UNICEF. 
Link: https://www.kaggle.com/ruchi798/malnutrition-across-the-globe

Sample of the malnourishment dataset:
```{r}
knitr::kable(malnourishment_sample, format = "html")
```

Sample of the health expenditures dataset:
```{r}
knitr::kable(health_expenditures_sample, format = "html")
```

The malnourishment dataset has a fairly simple and straightforward structure. It breaks down each set of statistics by country. The sample includes only a few of these statistics, separated into the columns “Wasting”, “Overweight”, and “Underweight”. These columns describe the percentage of children aged 0-59 months who meet the criteria for these conditions. 
1. Wasting: below -2 standard deviations from the median weight-for-height. 
2. Overweightness: above 2 standard deviations from the median weight-for-height. 
3. Underweightness: below -2 standard deviations from the median weight-for-age. 

The health expenditures dataset, just like the malnourishment dataset, breaks down each set of statistics by country. The statistics associated with each country are broken down by year in separate columns, and describe the spending on health in a country during a particular year, in the unit current US dollars. The columns in the sample, year 2017 and year 2018, indicate the years 2017 and 2018, respectively, and contain the health expenditures of a country in that year. 

## Section 2.2


```{r, echo = FALSE}
knitr::kable(overweight_density, format = "html")
```
This graph shows the distribution of the rates of overweightness by country. As can be seen on the graph, majority of countries have an overweight rate between 0 and 10%.

```{r, echo = FALSE}
knitr::kable(mal_effects_bar, format = "html")
```
This graph shows the global averages of rates of the different types of malnourishment in the dataset. Stunting is, by far, the most common form of malnourishment globally.

```{r, echo = FALSE}
knitr::kable(stunting_bubble, format = "html")
```
This graph displays the stunting rate percentage by income classification with 0 representing the poorest countries and 3 representing the richest countries. There is also each plot size that represents the population under 5 in each country. It can be seen that there is a declining correlation between the stunting prevalency and income classification.

```{r, echo = FALSE}
kable(knit_mal_df)
```

This shows the global averages of various malnourishment effects.

Bangladesh displays outliers in both stunting in the underweight category. It also has one of the largest under 5 populations in this dataset. South Sudan is also an outlier for the severe wasting and wasting categories. These could obviously skew the data in the respective categories but especially in the case of Bangladesh when weighting the population as it has one of the highest. The greatest outliers are within the population. The max is `r max_u5_pop` and the mean is `r mean_u5_pop` There is a portion of the data where the population is above 80,000 (by thousands) and then there is a large break down to below 30,000. In some evaluations this can alter the effects towards countries with larger populations and make false correlations that may be due to factors only present in those countries.
