# Section 2
The first dataset we will be using is a dataset of all the health expenditures per capita in each country. The second dataset we will be using includes the statistics on malnutrition amongst children aged 0-59 months, broken down by the four types of malnutrition: wasting, stunting, underweight, and overweight.

The first dataset comes from data prepared by the World Health Organization. “The health expenditure estimates have been prepared by the World Health Organization under the framework of the System of Health Accounts 2011 (SHA 2011). The Health SHA 2011 tracks all health spending in a given country over a defined period of time regardless of the entity or institution that financed and managed that spending. It generates consistent and comprehensive data on health spending in a country, which in turn can contribute to evidence-based policy-making.” (World Development Indicators)
Link: https://databank.worldbank.org/reports.aspx?source=2&series=SH.XPD.CHEX.PC.CD

The second dataset comes from data compiled by Ruchi Bhatia, originally sourced from UNICEF. 
Link: https://www.kaggle.com/ruchi798/malnutrition-across-the-globe

Sample of the malnourishment dataset:
```{r}
knitr::kable(malnourishment_sample, format = "html")
```

Sample of the health expenditures dataset:
```{r}
knitr::kable(health_expenditures_sample, format = "html")
```

The malnourishment dataset has a fairly simple and straightforward structure. It breaks down each set of statistics by country. The sample includes only a few of these statistics, separated into the columns “Wasting”, “Overweight”, and “Underweight”. These columns describe the percentage of children aged 0-59 months who meet the criteria for these conditions. 
1. Wasting: below -2 standard deviations from the median weight-for-height. 
2. Overweightness: above 2 standard deviations from the median weight-for-height. 
3. Underweightness: below -2 standard deviations from the median weight-for-age. 

The health expenditures dataset, just like the malnourishment dataset, breaks down each set of statistics by country. The statistics associated with each country are broken down by year in separate columns, and describe the spending on health in a country during a particular year, in the unit current US dollars. The columns in the sample, year 2017 and year 2018, indicate the years 2017 and 2018, respectively, and contain the health expenditures of a country in that year. 
