
# Section 3
### Data science question 1: 
What is the distribution of country counts across the range of income classifications? 

This question is examining the distribution of countries across the income classifications, which range from 0-3. This question seeks to determine which income classification has the highest number of countries, and which income classification has the lowest. 

My data analysis method starts with wrangling the data to only include each country as one observation (in the original dataframe, most countries are listed multiple times due to several different observation years). Because income classification remains the same across all these years for each country, I was able to simply remove duplicates of country names so that each country was only included as one observation. Next, I made the bar plot of counts using ggplot by including the built-in “stat” argument to have the height of the bar represent the count of cases within each income classification. 

```{r part_2, echo=FALSE}
plot(income_classification_graph)
```

Evaluation: 
The bar graph reveals that income classification “2” has the highest country count, around approximately 55, and income classification “3” has the lowest country count, at just above 20. It appears that the majority of countries lie in the middle of the distribution of income classifications, across classification “1” and “2”. 

### Data science question 2: 
What is the distribution of countries across the range of health expenditures? Does this distribution map onto the distribution of income classifications? 

This question is examining the distribution of countries across the range of health expenditures, which have a large range from the 100s to over 10,000. This question further seeks to examine whether this distribution correlates to the distribution of income classifications. 

My data analysis method starts with some simple data wrangling in order to remove duplicate observations of the same country and filter for the most recent year with data, which is 2018. I also had to change all the health expenditure values into continuous variables, because they were originally inputted as characters instead of numbers. Because every country had different health expenditures, I had to use vector binning to generalize the health expenditure values into smaller ranges. I chose to cut them into groups of 500, and created my bar graph off of these groups. 

```{r part_2, echo=FALSE}
plot(health_expenditures_graph)
```

Evaluation: 
The bar graph reveals that the distribution of health expenditures is skewed right, meaning the majority of values lie in lower health expenditure values. It appears that most countries have a health expenditure of 0-500. This does not map onto the distribution of income classifications, as the distribution of income classifications was close to symmetrical, most of the values being in the middle two income classifications. This is rather surprising because you would expect that income classification would be positively correlated with health expenditure, meaning the higher a country’s income classification, the higher their health expenditure. However, because these two bar graphs do not show the same distribution, it indicates that there is not a correlation between these features. 

### Data science question 3: 
Is there a correlation between countries' health expenditures and their average rate of overweightness for their population under 5?

This question is examining if there is any correlation between money spent on health and rates of overweightness in individuals under 5. Overweightness can be caused by both under and overnourishment.

I first took the dataset for malnourishment and omitted any NA values. Then I selected the only columns of interest (using the ISO.code to record the country). Next, I grouped by that ISO.code and found the mean of the overweight percentage and populations under 5 between the years 2011 and 2018. I, then, wrangled the health expenditure data to group by country, average the expenditures per country and join that data with my the malnourishment data. 

```{r part_2, echo=FALSE}
plot(overweight_expenditure_plot)
```

Evaluation:
After evaluation, I've determined there is no correlation between average overweightness and average health expenditures. I determined this quantitatively by running a correlation test on my data and checking the p-value. The p-value was 0.8755 which is very high and highly suggests there is no correlation in the data. This means there is an 88% chance that I would get data at least this extreme if the null hypothesis is true. I also found this graphically by plotting the data in a scatter bubble plot and also showing a regression line. The reason I performed this specific test was to maybe show an opposite correlation with what we have been previously seeing with other effects of malnourishment (lower expenditures results in more people effected). I figured overweightness may be more prevalent in richer countries due to surplus of food and supply. 
