## Section 2: Data Description

The first dataset we will be using is a dataset of all the health expenditures per capita in each country. The second dataset we will be using includes the statistics on malnutrition amongst children aged 0-59 months, broken down by the four types of malnutrition: wasting, stunting, underweight, and overweight.

The health expenditure data set is derived from data prepared by the World Health Organization. “The health expenditure estimates have been prepared by the World Health Organization under the framework of the System of Health Accounts 2011 (SHA 2011). The Health SHA 2011 tracks all health spending in a given country over a defined period of time regardless of the entity or institution that financed and managed that spending. It generates consistent and comprehensive data on health spending in a country, which in turn can contribute to evidence-based policy-making.” (World Development Indicators)
Link: https://databank.worldbank.org/reports.aspx?source=2&series=SH.XPD.CHEX.PC.CD

The second dataset comes from data compiled by Ruchi Bhatia, originally sourced from UNICEF. 
Link: https://www.kaggle.com/ruchi798/malnutrition-across-the-globe

**Sample of the malnourishment dataset:**
```{r malnourishment_sample, echo=FALSE, warning=FALSE}
kable(malnourishment_sample)
```

**Sample of the health expenditures dataset:**
```{r all_health_expenditures_sample_df, echo=FALSE, warning=FALSE}
kable(all_health_expenditures_sample_df, 
      col.names = c("Country", "2017", "2018"))
```

The malnourishment dataset has a fairly simple and straightforward structure. It breaks down each set of statistics by country. The sample includes only a few of these statistics, separated into the columns “Wasting”, “Overweight”, and “Underweight”, and "Stunting". These columns describe the percentage of children aged 0-59 months who meet the criteria for these conditions.<p>
1. Wasting: below -2 standard deviations from the median weight-for-height.<br> 
2. Overweightness: above 2 standard deviations from the median weight-for-height. <br>
3. Underweightness: below -2 standard deviations from the median weight-for-age.<br>
4. Stunting: below -2 standard deviations from median height-for-age. <br>

The health expenditures dataset, just like the malnourishment dataset, breaks down each set of statistics by country. The statistics associated with each country are broken down by year in separate columns, and describe the spending on health in a country during a particular year, in the unit current US dollars. The columns in the sample, year 2017 and year 2018, indicate the years 2017 and 2018, respectively, and contain the health expenditures (per capita) of a country in that year. 


### Section 2.2: Summary Analysis - Health Expenditure Data

**Summary Statistics: **

```{r summary_statistics, echo = FALSE, warning = FALSE}
kable(summary_statistics_df,
      col.names = c("", "mean", "max", "min"))
```


<p>The first row is the mean health expenditure per capita for all the countries in the world for the years of 2000, 2011, and 2018. The bottom two rows display the minimum and maximum health expenditures per capita of any country for these years. 


```{r expenditure_rate_change_plot, echo = FALSE, warning = FALSE}
plot(expenditure_rate_change_plot)
```

<p>According to this choropleth map, the majority of countries around the world saw a 0% to 700% increase in their health expenditures between the years of 2000 and 2018. The outliers in this data are Venezuela, the Democratic Republic of Congo, China, Myanmar, and Ecuador. The health expenditures of Venezuela and the Democratic Republic of Congo fell 27.65% and 0.59%, respectively. This is likely due to political turmoil, which weakened their health infrastructures. The health expenditures of China, Myanmar, and Ecuador rose 1,089.81%, 1,265.65%, and 977.94%, respectively. The most important factors that contributed to this are likely improving health and living standards. Since the data displays each country's statistics individually, outliers do not impact the results. 

```{r top_10_expenditures_plot, echo = FALSE, warning = FALSE}
plot(top_10_expenditures_plot)
```

<p>The bar graph reveals that the United States had the largest health expenditure per capita in 2018, followed by Switzerland and Norway. Since the data displays each country's statistics individually, outliers do not impact the result. 

```{r country_trend_plot, echo = FALSE, warning = FALSE}
plot(country_trend_plot)
```

<p>The line graph exposes the United States as being an outlier in regard to health expenditure trends between 2011 and 2018. While most of the countries' health expenditures on the graph fluctuated between 0 and 6,000 USD per capita, the United States saw a constant increase of health expenditure throughout these years (from 8130.76 USD to 10,623.85 USD). The magnitude and range of the increase of the health expenditure of the United States increases the average expenditure of the world, which is important to take into account in a world-level analysis.  

### Section 2.2: Summary Analysis - Malnourishment Data

**Summary Statistics:**

```{r knit_mal_df, echo = FALSE, warning=FALSE}
kable(knit_mal_df, col.names = c("Malnutrition Effect", "Global Average Rate"))
```

This shows the global averages of various malnourishment effects.
git brancg
```{r overweight_density, echo=FALSE, warning=FALSE}
plot(overweight_density)
```

This graph shows the distribution of the rates of overweightness by country. As can be seen on the graph, majority of countries have an overweight rate between 0 and 10%.

```{r malnourishment_categories_bar, echo = FALSE, warning=FALSE}
plot(malnourishment_categories_bar)
```

This graph shows the global averages of rates of the different types of malnourishment in the dataset. Stunting is, by far, the most common form of malnourishment globally.

```{r stunting_bubble, echo = FALSE, warning=FALSE}
plot(stunting_bubble)
```

This graph displays the stunting rate percentage by income classification with 0 representing the poorest countries and 3 representing the richest countries. There is also each plot size that represents the population under 5 in each country. It can be seen that there is a declining correlation between the stunting prevalency and income classification.


Bangladesh displays outliers in both stunting in the underweight category. It also has one of the largest under 5 populations in this dataset. South Sudan is also an outlier for the severe wasting and wasting categories. These could obviously skew the data in the respective categories but especially in the case of Bangladesh when weighting the population as it has one of the highest. The greatest outliers are within the population. The max is `r max_u5_pop` and the mean is `r mean_u5_pop`. There is a portion of the data where the population is above 80,000 (by thousands) and then there is a large break down to below 30,000. In some evaluations this can alter the effects towards countries with larger populations and make false correlations that may be due to factors only present in those countries.
